---
title: "Cyclistic Bike-Share Analysis"
author: "CJ"
date: "2024-12-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction

The purpose of this report is to analyze Cyclistic bike-share data to understand the differences between casual riders and annual members. This analysis will help inform marketing strategies for converting casual riders into annual members.

# Cyclistic Bike-Share Analysis

## Business Questions:
1. How do annual members and casual riders use Cyclistic bikes differently?
2. Why would casual riders buy Cyclistic annual memberships?
3. How can Cyclistic use digital media to influence casual riders to become members?

## Data Source:
The data used in this analysis is publicly available from the Divvy trip data, provided by Motivate International Inc.

## Methodology:
- Data cleaning and transformation were done using R.
- Visualizations were created with ggplot2.

## Key Findings and Recommendations:
1. Casual riders use bikes more frequently on weekends.
2. Annual members have a higher consistency in their usage.
3. Marketing campaigns should target high-traffic stations and peak weekends.

# Data Cleaning and Preprocessing

```{r}
# Load necessary libraries
library(tidyverse)

# Set working directory (adjust the path)
setwd("C:/LearnCodingLanguage/data-analytics-portfolio/Cyclistic-Bike-Share-Analysis/Divvy_Trips_All/data")

# Load data (example file)
file_list <- list.files(pattern = "*.csv")
all_data <- lapply(file_list, read_csv) %>% bind_rows()

# Clean data
all_data <- all_data %>%
  mutate(
    start_time = as.POSIXct(start_time, format = "%Y-%m-%d %H:%M:%S"),
    end_time = as.POSIXct(end_time, format = "%Y-%m-%d %H:%M:%S"),
    ride_length = as.numeric(difftime(end_time, start_time, units = "mins")),
    day_of_week = weekdays(as.Date(start_time))
  ) %>%
  filter(ride_length > 0)  # Remove invalid rides

# Check the first few rows of cleaned data
head(all_data)




# Calculate average ride length by user type
avg_ride_length <- all_data %>%
  group_by(usertype) %>%
  summarize(avg_length = mean(ride_length))

# Plot average ride length
ggplot(avg_ride_length, aes(x = usertype, y = avg_length, fill = usertype)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Ride Length by User Type", x = "User Type", y = "Average Ride Length (minutes)") +
  theme_minimal()

# Rides by day of the week
rides_by_day <- all_data %>%
  group_by(usertype, day_of_week) %>%
  summarize(total_rides = n())

# Plot rides by day of the week
ggplot(rides_by_day, aes(x = day_of_week, y = total_rides, fill = usertype)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Total Rides by Day of the Week", x = "Day of the Week", y = "Total Rides") +
  theme_minimal()
